Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-28 numerical-cooling.tex
Change a bit of wording and usages of macros.

As discussed in Section~\ref{sec.num.chemistry}, the set of reactions that
characterize a metal-free gas is simple enough to be computed in
non-equilibrium during even very simulations.  Similarly, the radiative
cooling of metal-free gas is solved by directly computing the cooling
and heating rates from the following individual processes for atomic H
and He: collisional excitation and ionization, recombination,
free-free emission, Compton scattering off of the cosmic microwave
background (CMB), and photo-heating from a UV metagalactic background.
If the H$_{2}$ chemistry network is enabled, the following H$_{2}$
cooling processes are also considered: ro-vibrational transitions
\citep{2008MNRAS.388.1627G,1998A&A...335..403G}, heating and cooling from molecular
formation and destruction \citep{2009Sci...325..601T}, and
collision-induced emission \citep{2004MNRAS.348.1019R}.  If Deuterium
chemistry is enabled, then rotational transitions of HD
\citep{1998A&A...335..403G, 1983ApJ...270..578L} are treated as well.  The
radiative cooling calculation is coupled to the update of the chemistry network
such that they both occur within the same subcycling loop.  This is necessary
in regimes where rapid cooling or change in the ionization state occur, as this
will influence the kinetic rate coefficients through both changes in energy and
the equation of state of the gas.  In addition to the subcycle timestepping
contraints mentioned in Section~\ref{sec.num.chemistry}, the subcycle timestep
is also not permitted to exceed 10\% of the cooling time, $t_{cool} =
e/\dot{e}$.  A metagalactic background affects the gas through both
photo-heating and photo-ionization.  These are treated by including
redshift-dependent photo-ionization and photo-heating rate terms in the
chemistry and cooling equations for \ion{H}{1}, \ion{He}{1}, and \ion{He}{2}.
The black curves in Figure \ref{fig.cooling_rate} show cooling rates for a
metal-free gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ both with
and without a radiation background.  More detail on the specific UV backgrounds
present in \enzo\ is presented below in Section~\ref{sec.num.rad-homogeneous}.

@@

A more sophisticated method makes use of multidimensional cooling and
heating rate tables computed with the photo-ionization code
\texttt{Cloudy} \citep{1998PASP..110..761F}.  This method, detailed in
\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}, works by
using \texttt{Cloudy} to compute the cooling and heating rates from
the metal species only.  The primary assumption made is that of
ionization equilibrium.  The tables can have up to five dimensions,
where those dimensions are density, metallicity, electron fraction,
redshift (for an evolving metagalactic UV background), and
temperature.  Tables can be created for any abundance pattern for
elements up to atomic number 30 (Zn) and for any incident radiation
field.  The cooling from H and He is calculated by the non-equilibrium
primordial chemistry and cooling solver.  The contribution of metals
to the cooling is computed within the same subcycling loop as the
coupled primordial chemistry and cooling solver.  The blue curves in
Figure \ref{fig.cooling_rate} show cooling rates calculated with this
method for a gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ and
metallicity of 0.5 Z$_{\odot}$.

>>>>

As discussed in Section~\ref{sec.num.chemistry}, the set of reactions that
characterize a metal-free gas is simple enough to be computed in
non-equilibrium during even very large simulations.  Similarly, the radiative
cooling of metal-free gas is solved by directly computing the cooling
and heating rates from the following individual processes for atomic H
and He: collisional excitation and ionization, recombination,
free-free emission, Compton scattering off of the cosmic microwave
background (CMB), and photo-heating from a UV metagalactic background.
If the \HH chemistry network is enabled, the following \HH
cooling processes are also considered: ro-vibrational transitions
\citep{2008MNRAS.388.1627G,1998A&A...335..403G}, heating and cooling from molecular
formation and destruction \citep{2009Sci...325..601T}, and
collision-induced emission \citep{2004MNRAS.348.1019R}.  If Deuterium
chemistry is enabled, then rotational transitions of HD
\citep{1998A&A...335..403G, 1983ApJ...270..578L} are treated as well.  The
radiative cooling calculation is coupled to the update of the chemistry network
such that they both occur within the same subcycling loop.  This is necessary
in regimes where rapid cooling or change in the ionization state occur, as this
will influence the chemical kinetic rate coefficients through both changes in
energy and the equation of state of the gas.  In addition to the subcycle
timestepping contraints mentioned in Section~\ref{sec.num.chemistry}, the
subcycle timestep is also not permitted to exceed 10\% of the cooling time,
$t_{cool} = e/\dot{e}$.  A metagalactic background affects the gas through both
photo-heating and photo-ionization.  These are treated by including
redshift-dependent photo-ionization and photo-heating rate terms in the
chemistry and cooling equations for \ion{H}{1}, \ion{He}{1}, and \ion{He}{2}.
The black curves in Figure \ref{fig.cooling_rate} show cooling rates for a
metal-free gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ both with
and without a radiation background.  More detail on the specific UV backgrounds
present in \enzo\ is presented below in Section~\ref{sec.num.rad-homogeneous}.

@@

A more sophisticated method makes use of multidimensional cooling and
heating rate tables computed with the photo-ionization code
\texttt{Cloudy} \citep{1998PASP..110..761F}.  This method, detailed in
\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}, works by
using \texttt{Cloudy} to compute the cooling and heating rates from
the metal species only.  The primary assumption made is that of
ionization equilibrium.  The tables can vary along up to five dimensions,
of density, metallicity, electron fraction,
redshift (for an evolving metagalactic UV background), and
temperature.  Tables can be created for any abundance pattern for
elements up to atomic number 30 (Zn) and for any incident radiation
field.  Cooling from the standard \enzo\ non-equilibrium cooling module is
applied on top of the metal contributions.  The contribution of metals
to the cooling is computed within the same subcycling loop as the
coupled primordial chemistry and cooling solver.  The blue curves in
Figure \ref{fig.cooling_rate} show cooling rates calculated with this
method for a gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ and
metallicity of 0.5 Z$_{\odot}$.
